\section{Existing Implementation}
In order to study the effects of impact ionisation in solar cells using small Hubbard clusters, a program was developed. The original numerical implementation was by Michael Innerberger \cite{innerberger}, and extensions were carried out by Paul Worm \cite{worm_bachelor, worm_project} and Paul Prauhart \cite{prauhart}.
\medskip

In this section we will only give a brief overview of the main functionality of the code and refer the reader back to the previously mentioned three papers for more details.

\subsection{Hubbard Model} \label{sec:hubbard_model}
The Hubbard model was used to study the effect of impact ionisation in a strongly correlated electron system . It describes a system of atoms as a lattice of $N_s\in \mathbb{N}$ sites, each of which can represent the location of at most one spin-up electron and one spin-down electron under the tight-binding approximation. The wave functions that describe such localised electrons are called Wannier-Functions.
\medskip

Using the second quantisation formalism of quantum mechanics, states of such many-body systems can be described purely by the occupation numbers of each site $n_{i\sigma} = \{0,1\}$. This allows us to write state vectors as $\ket{\psi}= \ket{n_{1\uparrow}n_{1\downarrow} n_{2\uparrow}n_{2\downarrow}\dots n_{N_{s}\uparrow}n_{N_{s}\downarrow}}$. The Hamiltonian can then be expressed using the following two terms
\begin{equation}
    \hat{H}_{\text{Hubbard}} = U \sum_i \hat{n}_{i\uparrow} \hat{n}_{i\downarrow} + \sum_{ij\sigma} v_{ji} \hat{c}^\dagger_{i\sigma} \hat{c}_{j\sigma}\label{eq:hubbard_hamiltonian}
\end{equation}
The first corresponds to the repulsive Coulomb interaction $U$ between two electrons with opposite spins located on the same site. The second represents the energy change in the system when an electron hops from site $j\to i$ with hopping amplitude $v_{ji}$. Here $\hat{c}_{i\sigma}^\dagger, \hat{c}_{j\sigma}$ are the fermionic creation and annihilation operators.
\medskip

One can interpret the first sum as representing the interaction energy (also sometimes referred to as potential energy) in the system, whereas the second sum for the hoppings represents the system's kinetic energy since it is related to the movement of electrons. This distinction becomes vital in section \ref{sec:non_local_coulomb}, where we introduce non-local Coulomb interactions.



\subsection{Interaction with Photons}
The systems interaction with photons is modelled via a classical electric field pulse
\begin{equation}
    \Vec{E}(t) = \Vec{E}_0 \sin(\omega(t-t_p))\operatorname{e}^{-\frac{(t-t_p)^2}{2\sigma^2}} \label{eq:e_field}
\end{equation}
with frequency $\omega$, width $\sigma$ and peak time $t_p$. This can be integrated into our model by introducing a time-dependent phase factor onto the hopping amplitude in a method called Peierl's substitution \cite{peierl}.
\begin{equation}
    v_{ij} \to v_{ij}(t) = v_{ij}\exp\left(i\frac{e}{\hbar} \int_{\Vec{R}_i}^{\Vec{R}_j} \Vec{A}(\Vec{r},t) d\Vec{r}\right)\label{eq:hopping}
\end{equation}
Here $\Vec{A}$ is the electromagnetic vector potential which, in a gauge where the scalar potential vanishes, can be expressed via $\Vec{E}(t) = -\partial_t \Vec{A}(t)$. Using an approximation for the integral in (\ref{eq:hopping}) and the equation (\ref{eq:e_field}), one arrives at
\begin{equation}
    v_{ij} \approx v_{ij} \exp\left(ia[\cos(\omega (t-t_p))-b] e^{-\frac{(t-t_p)^2}{2\sigma^2}}\right)
\end{equation}
where $a$ and $b$ are tunable parameters.


\subsection{Time-Evolution}

Our main interest is to investigate impact ionisation in the system after being exposed to the electric field pulse. Prior to this work, it was achieved by looking at the double occupation observable $\braket{\hat{d}(t)} = \braket{\hat{n}_{i\uparrow}(t) \hat{n}_{i\downarrow}(t)}$ since the rise of its mean value after initial excitation was an indicator for impact ionisation. In order to compute this quantity, we need the ability to time-evolve any initial state $\ket{\psi(t=0)} = \ket{\psi_0}$ of the system. Using the time-dependent Schrödinger equation
\begin{equation}
    i\hbar \frac{\partial}{\partial t}\ket{\psi(t)} = \hat{H}(t)\ket{\psi(t)}\label{eq:time_evolve}
\end{equation}
the time evolution can be computed as
\begin{equation}
    \ket{\psi (t)} = \mathcal{T} \exp\left(-\frac{i}{\hbar}\int_0^t H(t') dt'\right) \ket{\psi_0}
\end{equation}
where $\mathcal{T}$ is the time-ordering operator. By numerically dividing $t$ into $m$ small time steps $\tau$, it is justified to use Magnus-expansion \cite{magnus} of order zero and thus neglect the time-ordering operator. Assuming the time steps are small enough, we can approximate the integral in (\ref{eq:time_evolve}) over one time step using the midpoint rule. This leads to the following recursive formula
\begin{equation}
    \ket{\psi(m\tau + \tau)} = \exp\left(-\frac{i}{\hbar} H\left(m\tau + \frac{\tau}{2}\right)\right)\ket{\psi(m\tau)}.
\end{equation}
This can be computed using the Krylov matrix exponential method \cite{innerberger}.


\subsection{Finding the Initial State}
We mainly concern ourselves with half-filled systems when studying impact ionisation, where the number of spin-up and spin-down electrons are the same. The number of electrons is a conserved property of the systems and does not change over time.
\medskip

It is assumed that before the electric field pulse, the system is in thermal equilibrium and occupies the state with the lowest energy (the ground state). To find this state numerically, a variant of the power iteration method was implemented \cite{innerberger}. From a randomly initialised starting state, it iteratively computes the eigenenergy with the largest absolute value and its corresponding eigenstate, thus recursively converging to the ground state.


\subsection{Memory Management}
The predominant limitation of this implementation is the memory it requires. For a system with $N_s$ sites and a fixed number of electrons, the dimension of its Hilbert space, i.e. the number of linearly independent states, is given by 
\begin{equation}
    \dim \left[\mathcal{H}^{n_\uparrow}_{n_\downarrow} (N_s)\right] = \begin{pmatrix}N_s \\ n_\uparrow\end{pmatrix} \begin{pmatrix}N_s \\ n_\downarrow\end{pmatrix}
\end{equation}
which for a half-filled system with $14$ sites is about $\SI{12e6}{}$. A Hamiltonian matrix of that size, with elements of the data type double, would take over $1000$ Terabytes of memory. However, due to the Hamiltonian being made up of mostly zeros, it can be stored and manipulated in a highly compressed sparse matrix format \cite{innerberger}, substantially reducing the memory needs.
\medskip

The states of the system are stored as integers whose binary representations correspond to the occupation configuration of the sites. Actions on these states like creation, annihilation and hoppings have been implemented as bitwise operations.
\medskip
 
With all these data-saving measures in place, it was possible to compute the time evolution of 2D square lattices and chains of up to 14 sites, above which memory becomes the limiting factor once again.


\subsection{Spectral Functions}\label{sec:spectral_functions}

One way of obtaining information about the states of a system is via spectral functions $A(\omega, t)$. They provide insight into which states and energies the electrons in the system can occupy and thus also tell us about the energies required for optical electron transitions to occur. The existing codebase offers two methods of computing spectral functions \cite{spectral_function}:

\subsubsection{Lehmann Spectra}
Implemented by M. Innerberger in \cite{innerberger}, the Lehmann representation of the spectral function gives the equilibrium, i.e. time-independent, spectrum of the system. 
\begin{equation}
    A(E)=\sum_{i, \sigma} \sum_{|\phi\rangle}\left(\left|\left\langle\phi\left|\hat{c}_{i \sigma}^{\dagger}\right| \psi_{0}\right\rangle\right|^{2} \delta\left(E-E_{|\phi\rangle}+E_{0}\right)+\left|\left\langle\phi\left|\hat{c}_{i \sigma}\right| \psi_{0}\right\rangle\right|^{2} \delta\left(E+E_{|\phi\rangle}-E_{0}\right)\right)
\end{equation}
Where $\{ \ket{\phi}\}$ is an eigenbasis of $\mathcal{H}(N_s)$ with respective energy $E_{\ket{\phi}}$, and $\ket{\psi_0}$ is the ground state of $\hat{H}$ with energy $E_0$.

\bigskip
It is quick and straightforward to generate with the existing code and produces numerically elegant results. However, because it is pretty memory intensive, it can not handle systems with a large number of sites. This, together with the fact that we can not investigate the time dependence of the spectrum, prompted the implementation of another way to calculate the non-equilibrium spectral function.

\subsubsection{Fourier Transforms of Non-Equilibrium Green's Functions}

Initially implemented by P. Worm \cite{worm_bachelor}, with efficiency improvements by P. Prauhart \cite{prauhart}, the one-particle non-equilibrium greens function $G_{ij\sigma} (t,t')$ can be used to compute the non-equilibrium spectral function $A_{ij\sigma}(\omega, t)$, via forward Fourier transformation.

\begin{equation}
    G_{i j \sigma}^{<}\left(t, t^{\prime}\right)=\mathrm{i}\left\langle\psi\left(t^{\prime}\right)\left|\hat{c}_{j \sigma}^{\dagger} \mathcal{T} e^{-\mathrm{i} \int_{t}^{t^{\prime}} H(\tau) \mathrm{d} \tau} \hat{c}_{i \sigma}\right| \psi(t)\right\rangle\label{eq:greens_function_lesser}
\end{equation}

\begin{equation}
    G_{i j \sigma}^{>}\left(t, t^{\prime}\right)=\mathrm{i}\left\langle\psi\left(t^{\prime}\right)\left|\hat{c}_{j \sigma} \mathcal{T} e^{-\mathrm{i} \int_{t}^{t^{\prime}} H(\tau) \mathrm{d} \tau} \hat{c}_{i \sigma}^\dagger\right| \psi(t)\right\rangle\label{eq:greens_function_greater}
\end{equation}

\begin{equation}
    A_{ij\sigma}^\gtlt = \frac{1}{\pi} \operatorname{Im}\int_0^\infty \operatorname{e}^{i\omega t_{rel}} G_{ij\sigma}^\gtlt (t,t+t_{rel}) dt_{rel}\label{eq:non_equilibirum_spectral_function}
\end{equation}

A simple interpretation of the definition in (\ref{eq:greens_function_lesser}) is that an electron with spin $\sigma$ is removed from site $i$ of the system at time $t$, the system is time evolved to $t'$, and then the electron is added back at site $j$. $G_{ij\sigma}^{<}$ is then proportional to the probability that this system is in the same state as a system at time $t'$ where the electron was never removed. The exact implementation is elaborated on in \cite{worm_bachelor}. 

\subsubsection{Obtaining spectral functions using the existing code}

In the current state of the code, the Lehmann spectrum can be generated automatically and is output as a .dat file containing columns of values of $A(\omega)$ for each site.

\bigskip
The Green's function is also output as a .dat file with columns containing complex number values of $G(t,t'=t_0)$ for each site. In order to obtain the spectral function from this, one needs to perform the Fourier transform of this data manually, using (\ref{eq:non_equilibirum_spectral_function}). There is currently no existing function in the codebase that handles this automatically. In order to obtain more usable, less noisy results, one should, prior to the Fourier transform, multiply $G$ with a dampening factor $e^{-\epsilon (t-t_0)}$, which effectively acts as a Lorentzian broadening in frequency space \cite{spectral_function}. To achieve a higher frequency/energy resolution, one can additionally zero-pad the domain before performing the Fourier transform. 